### 2011-05-27 thesis/contents.tex
Added a lot of content from the techreport and some more structed and ideas.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Important sections of this chapter begin here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probability Theory}
\subsection{Bayes Theorem}

\section{Probabilistic Graphical Models}
\label{sec:graphical-models}
\subsection{Factor Graphs}
\subsection{Inference Engines}
Exact Inference and Approximate Inference methods

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Novelty Detection}\label{chap:novelty-intro}
\section{Novelty Detection as a Threshold}
\subsection{Conditional Probability}
\subsection{Unconditional Probability}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Novelty Detection on Semantic Representations}\label{chap:novelty}
\section{Detecting New Classes} 
\section{Detecting New Structure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Important sections of this chapter begin here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{Probability Theory}
\subsection{Bayes Theorem}

\section{Probabilistic Graphical Models}
\label{sec:graphical-models}
% THIS IS ALL JUNK TEXT COPIED
Graphical models usage can be tracked back to earlies 1920 but they only become
popular in mid-eighties when researchers started to use \emph{Bayesian Networks}
to model expert systems~\citep{borgelt2002graphical}.

They serve as a better tool to model \emph{random variables}
(nodes on the graph) and their probabilities as they model the conditional
dependence between variables (edges on the graph). Important to note that here
\emph{random variables} does not denotes a truly random variable but one that is
unknown by the system and is conditioned by other variables/evidences.

This type of graphs provide a generative model where the probability of any
given scenario can be determined. As opposed to non-generative models where a
given probability can only be calculated if certain data is given.
This means when a graphical model is learned, it can be used to generate new
samples from the learned distribution.

They have been successfully used in several machine-learning task such as:
information extraction, speech recognition, computer vision.
They are also useful due to their ability to deal with semantic (high-level)
features~\citep{boutell2006factor} and ability to represent properties of the
reality they try to model.

Two main types of graphical models have been used: Bayesian Networks which model
directed edges between variables and Markov Random Fields where variables are
connected by a potential but no special direction is given to edges.

An important property of these graphs is the \emph{Markov-blanket} of a node.
For a given variable $a$, a \emph{Markov-blanket} is a set of variables in the
surroundings of $a$ that when given the value of $a$ becomes independent of the
rest of the graph~\citep{pearl1988probabilistic}.
On non-directional graphs it is directly determined by the nodes connected to $a$.
This allows the usage of graph algorithms such as \emph{min-cut} to quickly
determine most likely scenarios. In the case of directed edges a node blanket is
also influenced by the direction of the edges and a more complex schemes need to
be used.

As \cite{lauritzen2002chain} points this two types of models can be represented
as a \emph{chain graph} where both directed and undirected edges can co-exist.
Though this generalization is hard to implement due to mis-understandings on the
concepts the graph-models use.

Another useful interpretation of graphical-models are \emph{factor graphs}.
Those are able to handle both \emph{Bayesian Networks} and
\emph{Markov Random Fields}. Under this interpretation a graph is seen as a
bipartite graph that connects variables with factors that influence
them~\citep{bishop2006pattern}.
This gives a very useful framework to develop belief propagation on them by
seeing a message-passing mechanism between nodes. Belief propagation is used to
calculate marginal-probabilities.

\begin{figure}[ht]
    \begin{minipage}[b]{0.3\linewidth}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/graphical-models/BayesNet.pdf}
        \caption{Bayes Networks}
    \end{minipage}
    \hspace{0.5cm}
    \begin{minipage}[b]{0.3\linewidth}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/graphical-models/MarkovRandomField.pdf}
        \caption{Markov Random Field}
    \end{minipage}
    \hspace{0.5cm}
    \begin{minipage}[b]{0.3\linewidth}
        \centering
        \includegraphics[width=\textwidth]{figures/graphical-models/FactorGraph.pdf}
        \caption{Factor Graph}
    \end{minipage}
\end{figure}

\subsection{Factor Graphs}
A \emph{factor graph} is a bipartite graph connecting two sets of nodes $X_G$ and $F_G$
representing random variables and factors.
Each factor is described by function $\phi$ dependent only on the variables $x_\phi$
to which the factor is connected.
Thus, a factor graph can be seen as a description of probability density function obtained
by a product of all the factors. In order to represent the probability,
a normalization factor needs to be introduced, resulting in the following equation:

\begin{equation}
P_G(x) = \frac{1}{Z}\prod_{\phi \in F_G}{\phi(x_{\phi})},\qquad
Z = \sum_{X_G}\prod_{\phi \in F_G}{\phi(x_{\phi})}
\end{equation}

More often we are only interested in a marginal of the probability distribution.

\begin{equation}
M_G(x) = \frac{1}{Z}\sum_{X_G \backslash x}{\prod_{\phi \in F_G}{\phi(x_{\phi})}}
\end{equation}

\subsection{Inference Engines}
Exact Inference and Approximate Inference methods

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Novelty Detection}\label{chap:novelty-intro}
\begin{itemize}
\item Introduction
\item History
\item Approaches to Novelty Detection
\end{itemize}
\section{Novelty Detection Review and Related Work}

\section{Novelty Detection as a Threshold}
When dealing with a statistic point of view, noisy data or unstable features, a decision to
classify a sample as novel has an associated error probability of $P(\overline{novel}|x)$.
And the error rate of a system that classifies the set of inputs $N$ as novel is given by
the equation:

\begin{equation}
P(e) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
\end{equation}

Therefore any novelty system that decides to acknowledge a sample
$a$ as novel should also detect novelty on any sample $b$ satisfying
$P(\overline{novel}|b) < P(\overline{novel}|a)$.
As that would always reduce the error probability.
This way a novelty detection system is interested in defining an
order relation on all the possible inputs equivalent to the order defined by
the error rate: $P(\overline{novel}|x)$.

Using Bayes rule and assuming a constant $P(\overline{novel})$
a ratio between a conditional and unconditional probabilities of the input $x$ is obtained.
Such a ratio is a suitable function for implementing a novelty detector system with optimal
thresholding.

\begin{equation}
\label{eq:novelty-threshold}
          P(\overline{novel}|x)
  =       \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)}
  \propto \frac{P(x|\overline{novel})}{P(x)}
\end{equation}

\subsection{Conditional Probability}
\subsection{Unconditional Probability}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Novelty Detection on Semantic Representations}\label{chap:novelty}
On the previous chapter, novelty as been introduced as a statistical
measure on the error-likelihood ratio between a conditional and unconditional
probability of a given .

\section{Detecting New Classes}
Explain exactly and formally how to obtain a novelty measure on a single
variable of a graph.\footnote{I am sure on the math of this.}

\section{Copping with Multiple Novel Variables}
On the previous section we show how to detect a novel class but that assumes
the rest of the graph holds true.

In this section we show how to handle novelty in several variables
in simultaneous.\footnote{I am almost sure the math here will work out the
way I think}

\subsection{Handling novelty signals from the low-level classifiers}
Explain how to handle novelty signals from low-level classifiers, now that we
have a model with the ability to still reason when several variables can be
novel.

\section{Detecting New Structure}
